### 2016-11-06 Paper_v1.tex
Nov6

% Load packages
\usepackage{url}  % Formatting web addresses
\usepackage{ifthen}  % Conditional
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
\usepackage{setspace}
%\usepackage{longtable}
\usepackage{colortbl}
%\usepackage{palatino,lettrine}
%\usepackage{times}
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage[wide]{sidecap}
%\usepackage[authoryear,round,comma,sort&compress]{natbib}
\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
%\usepackage[authoryear,round]{natbib}
\usepackage{supertabular}
\usepackage{simplemargins}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{lineno}
%\usepackage{chicago}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{amsmath}
%\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
\DeclareMathOperator*{\argmin}{\arg\!\min}

@@

\subsection*{Maximum theoretical CAT yield showed CFPS can be optimized.}
We calculated the carbon yield of CAT production for our experimental data and our best-fit parameter set as a function of the initial and final concentrations and the carbon numbers of CAT, glucose, and amino acids.
Arginine and glutamate were excluded due to not being present in the training dataset.
The experimental data displayed a CAT yield of 0.0865, while the best-fit parameter set displayed a CAT yield of 0.0871.
We then used sequence-specific FBA to calculate a theoretical maximum CAT yield of 0.1942.
Thus, we showed that our experimental dataset and best-fit parameter set were each producing CAT at 45\% of the theoretical maximum.
This allowed us to quantify the amount of carbon being diverted to byproducts, and suggests that there is potential for a doubling of CAT production by reducing this diversion of carbon.

@@

\subsection*{Calculation of CAT yield}
The yield on CAT production was calculated for three cases: the experimental data, the best-fit parameter set, and a theoretical maximum yield.
In each case the yield was formulated as a ratio of carbon produced as CAT to carbon consumed as reactants (glucose and amino acids):
\begin{equation}\label{eqn:yield-definition}
	Yield=\frac{\Delta CAT\;C_{CAT}}{\sum_{i=1}^{\mathcal{R}}\max(\Delta m_{i},0)\;C_{m_i}}
\end{equation}
where $\Delta CAT$ denotes the amount of CAT produced, $C_{CAT}$ denotes carbon number of CAT, $\mathcal{R}$ denotes the number of reactants, $\Delta m_{i}$ denotes the amount of the $i$th reactant consumed, never allowed to be negative, and $C_{m_i}$ denotes the carbon number of the $i$th reactant.
Because no data was available for arginine or glutamate, these reactants were left out of all three calculations.
In the experimental case and the best-fit set case, yield was calculated by setting $\Delta CAT$ equal to the final minus the initial CAT concentration and setting $\Delta m_{i}$ equal to the initial minus the final reactant concentration.
The theoretical yield was calculated using flux balance anaylsis (FBA) with a sequence-based analysis on CAT.
The sequence specific FBA \cite{2002_allen_palsson} problem was formulated as:
\begin{equation}\label{eqn:FBA}
\begin{split}
	\max_{\boldsymbol{v}}{} \! \left( v_{obj}=\mathbf{\boldsymbol{c}}^T \boldsymbol{v} \right) \qquad \alpha_i \leq v_i \leq \beta_i  \qquad i=1,2,\hdots,\mathcal{R} \\
	\mathrm{Subject \; to:}	\; \; \mathbf{S}\mathbf{v}=\mathbf{0}\qquad\qquad\qquad\qquad\qquad\qquad\qquad\quad
\end{split}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{v}$ denotes the unknown flux vector, $\boldsymbol{c}$ denotes the objective selection vector, and $\alpha_i$ and $\beta_i$ denote the lower and upper bounds on flux $v_{i}$, respectively.
The stoichiometric matrix was expanded to include the transcription and translation reactions for producing CAT.
The objective $v_{obj}$ was to maximize the specific rate of CAT formation.
The specific glucose uptake rate was constrained to allow a maximum flux of 10 mM/hr \cite{2002_Mahadevan_BiophysJ}; the amino acids and oxygen uptake rates were also bound to allow a maximum flux of 10 mM/hr, but did not reach this maximum flux.
Glucose, oxygen, and amino acids were modeled as being imported into the system, whereas CAT synthesis was modeled through protein export.
The rest of the network followed a psuedo steady-state asusmption where all other metabolites were not allowed to accumulate; thus, the network could be solved by linear programming.
The flux balance analysis problem was solved using the GNU Linear Programming Kit (v4.52) \cite{GLPK}.
The solution flux vector was used to calculate the theoretical carbon yield of CAT, reformulated in terms of flux:
\begin{equation}\label{eqn:yield}
	Yield=\frac{v_{CAT}\;C_{CAT}}{\sum_{i=1}^{\mathcal{R}}\max(v_{i},0)\;C_{i}}
\end{equation}
where $v_{CAT}$ denotes the CAT export flux and $v_{i}$ denotes the import flux of the $i$th substrate.

@@

% Figures and captions go here ...
\begin{figure}[ht]
\centering
\includegraphics[width=0.75\textwidth]{./Figures/Network.pdf}
\caption{Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, and NADPH/NADH transfer. FBA flux value (top), and mean ± standard error across ensemble at 1.5 hrs (middle) and 3 hrs (bottom), normalized to CAT synthesis flux.}
\label{fig:Network}
\end{figure}

\begin{figure}[ht]
\centering
%\includegraphics[width=1.00\textwidth]{./Figures/BothCarbon.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/Carbon.pdf}
\includegraphics[width=0.05\textwidth]{./Figures/null.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/CarbonNoControl.pdf}
\caption{Central carbon metabolism in the presence (top) and absence (bottom) of allosteric control, including glucose (substrate), CAT (product), and intermediates, as well as total concentration of energy species. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 18,000 sets.}
\label{fig:BothCarbon}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Energy.pdf}
\caption{Energy species and energy totals by base in the presence of allosteric control. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 18,000 sets.}
\label{fig:Energy}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Amino.pdf}
\caption{Amino acids in the presence of allosteric control. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 18,000 sets.}
\label{fig:Amino}
\end{figure}

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.94\textwidth]{./Figures/ClusterCAT.pdf}
\caption{Normalized first-order and pairwise sensitivities of CAT production to rate constants. Dark pixels indicate the most sensitive indices. Dendrogram indicates high responders (dark gray background), medium responders (medium gray background), and low responders (light gray background).}
\label{fig:ClusterCAT}
\end{figure}

@@

%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.00\textwidth]{./Figures/MikeYield.pdf}
%\caption{}
%\label{fig:MikeYield.pdf}
%\end{figure}

>>>>

% Load packages
\usepackage{url}  % Formatting web addresses
\usepackage{ifthen}  % Conditional
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{mathtools}
\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
\usepackage{setspace}
%\usepackage{longtable}
\usepackage{colortbl}
%\usepackage{palatino,lettrine}
%\usepackage{times}
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage[wide]{sidecap}
%\usepackage[authoryear,round,comma,sort&compress]{natbib}
\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
%\usepackage[authoryear,round]{natbib}
\usepackage{supertabular}
\usepackage{simplemargins}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{lineno}
%\usepackage{chicago}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{amsmath}
%\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
\DeclareMathOperator*{\argmin}{\arg\!\min}

@@

\subsection*{Maximum theoretical CAT yield showed CFPS can be optimized.}
We calculated the carbon yield of CAT production for our experimental data and our best-fit parameter set as a function of the initial and final concentrations and the carbon numbers of CAT, glucose, and amino acids.
%Arginine and glutamate were excluded due to not being present in the training dataset.
The experimental data displayed a CAT yield of 0.0865, while the best-fit parameter set displayed a CAT yield of 0.0871.
While the model ensemble described the experimental data, it was unclear whether the performance of the CFPS system was optimal.
To address this question, we used ssFBA in combination with the cell-free metabolic network and a detailed promoter model under a T7 polymerase to compute the maximum theoretical carbon yield.
However, we first validated the ssFBA approach by comparing an ensemble of simulated versus measured concentrations of CAT over a one hour period (Fig. ~\ref{fig:CATProdYield}A). 
The ensemble of 100 sets captured the CAT concentration profile which was randomly generated by sampling RNA polymerase levels, ribosome levels and elongation rates in a physiological range.
We then used sequence-specific FBA to calculate a theoretical maximum CAT yield under four different cases: unconstrained, limited oxidative phosphorylation, bounded by transcription/translation rates, and bounded by experimental data (Fig. ~\ref{fig:CATProdYield}B).
The theoretical maximum carbon yield of CAT was 0.349 for an unconstrained case and 0.194 for the transcription and translation constrained case.
Thus, we showed that our experimental dataset and best-fit parameter set were each producing CAT at 25\% of the theoretical maximum and 45\% of a theoretical physiological case.
Whereas, the case constrained by experimental data had a carbon yield of 0.062 $\pm$ 0.02, similar to the experimental yield. 
This allowed us to quantify the amount of carbon being diverted to byproducts, and suggests that there is potential for a doubling of CAT production by reducing this diversion of carbon.
In comparing the flux distributions between the unconstrained and constrained cases (Fig. ~\ref{fig:Network}), both constrained cases heavily utilize the Entner–Doudoroff pathway which may be a first viable knockout to increase CAT yield.

@@

\subsection*{Sequence specific FBA and calculation of CAT yield}
The yield on CAT production was calculated for each case as a ratio of carbon produced as CAT to carbon consumed as reactants (glucose and amino acids):
\begin{equation}\label{eqn:yield-definition}
	Yield=\frac{\Delta CAT\;C_{CAT}}{\sum_{i=1}^{\mathcal{R}}\max(\Delta m_{i},0)\;C_{m_i}}
\end{equation}
where $\Delta CAT$ denotes the amount of CAT produced, $C_{CAT}$ denotes carbon number of CAT, $\mathcal{R}$ denotes the number of reactants, $\Delta m_{i}$ denotes the amount of the $i$th reactant consumed, never allowed to be negative, and $C_{m_i}$ denotes the carbon number of the $i$th reactant.
Because no data was available for arginine or glutamate, these reactants were left out of all three calculations.
In the experimental case and the best-fit set case, yield was calculated by setting $\Delta CAT$ equal to the final minus the initial CAT concentration and setting $\Delta m_{i}$ equal to the initial minus the final reactant concentration.
The theoretical yield was calculated using flux balance anaylsis (FBA) with a sequence-specific based analysis on CAT.
The sequence specific FBA \cite{2002_allen_palsson} problem was formulated as:
\begin{equation}\nonumber
 \begin{multlined}
	\qquad \qquad \qquad \max_{\boldsymbol{w}}{} \! \left( w_{obj} = \mathbf{\boldsymbol{\theta}}^T \boldsymbol{w} \right) \\
	\mathrm{Subject \; to:}
	 \; \; \mathbf{S}\mathbf{w}=\mathbf{0} \\
\alpha_i \leq w_i \leq \beta_i  \qquad i=1,2,\hdots,\mathcal{R}
 \end{multlined}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{w}$ denotes the unknown flux vector, $\boldsymbol{\theta}$ denotes the objective selection vector
and $\alpha_i$ and $\beta_i$ denote the lower and upper bounds on flux $w_{i}$, respectively.
The objective $w_{obj}$ was to maximize the specific rate of CAT formation.
The specific glucose uptake rate was constrained to allow a maximum flux of 40 mM/hr according to literature data; the specific amino acid uptake rates were also bound to allow a maximum flux of 30 mM/hr, but did not reach this maximum flux.
The transcription and translation template reactions were added to the metabolic network and are based off sequence specific analysis \cite{2002_allen_palsson} involving transcription initiaion, transcription, mRNA degradation, translation initiation, translation, and tRNA charging. 
The mRNA and protein sequence of each protein was determined from literature. 
The transcription rate was constrained as:
\begin{equation}\nonumber
	w_{tx} = RNAP\frac{v_{RNAP}}{l_{mRNA}}\left(\frac{Gene}{km+Gene}\right)P
\end{equation}
where $RNAP$ is the concentration of RNA polymerase, $v_{RNAP}$ is the elongation rate (nucleotides/hr) by the RNA polymerase, $l_{mRNA}$ is the number of nucleotides in the mRNA, $Gene$ is the gene concentration, $km$ is the plasmid saturation coefficient, and $P$ is the promoter activity.

The promoter activity was formulated following Moon et al. for synthetic circuits as:
\begin{equation}\nonumber
	P = \frac{K_{1}}{1 + K_{1}}
\end{equation}
where $K_{1}$ represents the state of T7 RNA polymerase binding.

The translation rate was constrained as:
 \begin{equation}\nonumber
	w_{tl} = K_{P}Ribo\frac{v_{Ribo}}{l_{protein}}[mRNA_{ss}]
\end{equation}
where $K_{P}$ is the polysome amplification constant, $Ribo$ is the ribosome concentration,  $v_{Ribo}$ is the elongation rate (amino acids/hr) of the ribosome, $l_{protein}$ is the number of amino acids in the protein of interest, and $mRNA_{ss}$ is the mRNA concentration at steady state determined by the transcription rate divided by the degradation rate of mRNA.  
An ensemble of flux distributions was calculated for 100 sets by randonmly sampling. 
Glucose, oxygen, and amino acids were modeled as being imported into the system, whereas CAT synthesis and metabolite byproduct formation was modeled as an export from the system.
The rest of the network followed a pseudo steady-state asusmption where all other metabolites were not allowed to accumulate; thus, the network could be solved by linear programming.
The flux balance analysis problem was solved using the GNU Linear Programming Kit (v4.52) \cite{GLPK}.
The solution flux vector was used to calculate the theoretical carbon yield of CAT for four different cases.
%\begin{equation}\label{eqn:yield}
%	Yield=\frac{w_{CAT}\;C_{CAT}}{\sum_{i=1}^{\mathcal{R}}\max(w_{i},0)\;C_{i}}
%\end{equation}
%where $w_{CAT}$ denotes the CAT export flux and $w_{i}$ denotes the import flux of the $i$th substrate.
For the unconstrained case, all rates were left unbounded. 
An ensemble of flux distributions was calculated by randonmly sampling the maximum specific glucose uptake rate from 30 to 40 mM/hr determined from experimental data.
For the limited oxidative phosphorylation case, an ensemble of flux distributions was generated by randomly sampling the maximum specific oxygen uptake rate to from 0 to 10 mM/hr.
For the TXTL case (constrained by transcription and translation rates), an ensemble was generated by randomly sampling RNAP polymerase levels, ribosome levels, and elongation rates in a physiological range determined from literature as well as the specific glucose uptake rate from 30 to 40 mM/hr. 
For the case where the flux was constrained by experimental data, the lower and upper bounds where data was avaible were randomly sampled in the physiological range of the data in addition to randomly sampling RNAP polymerase levels, ribosome levels, and elongation rates in a physiological range determined from literature. 
\clearpage

@@

% Figures and captions go here ...
\begin{figure}[ht]
\centering
\includegraphics[width=0.75\textwidth]{./Figures/CAT_prod_yield.pdf}
\caption{Sequence specific flux balance analysis of CAT production and yield. A. 95\% confidence interval of the ensemble (light blue region) for CAT concentration versus time. B. Theoretical maximum carbon yield of CAT calcualted by ssFBA for four different cases: unconstrained (black), limited oxidative phosphorylation activity (gray), constrained by transcription and translation rates (light grey), and constrained by experimental data (white). Error bars represent standard deviation of the ensemble.   }
\label{fig:CATProdYield}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth]{./Figures/CAT_flux_final.pdf}
\caption{Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, NADPH/NADH transfer, and oxidative phosphorylation. Sequence specific FBA flux value (mean ± standard deviation) across ensemble for 1 hr, normalized to glucose uptake flux. Unconstrained (top row), constrained by transcription and translation rates (second row), and constrained by experimental data (bottom row) shows the flux distribution throughout central carbon metabolism.}
\label{fig:Network}
\end{figure}

\begin{figure}[ht]
\centering
%\includegraphics[width=1.00\textwidth]{./Figures/BothCarbon.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/Carbon.pdf}
\includegraphics[width=0.05\textwidth]{./Figures/null.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/CarbonNoControl.pdf}
\caption{Central carbon metabolism in the presence (top) and absence (bottom) of allosteric control, including glucose (substrate), CAT (product), and intermediates, as well as total concentration of energy species. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 100 sets.}
\label{fig:BothCarbon}
\end{figure}

\begin{figure}[ht]
\centering
%\includegraphics[width=1.00\textwidth]{./Figures/BothCarbon.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/Carbon.png}
\includegraphics[width=0.05\textwidth]{./Figures/null.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/CarbonNoControl.png}
\caption{Central carbon metabolism in the presence (top) and absence (bottom) of allosteric control, including glucose (substrate), CAT (product), and intermediates, as well as total concentration of energy species. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 100 sets.}
\label{fig:BothCarbon_png}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Energy.pdf}
\caption{Energy species and energy totals by base in the presence of allosteric control. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 100 sets.}
\label{fig:Energy}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Energy.png}
\caption{Energy species and energy totals by base in the presence of allosteric control. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 100 sets.}
\label{fig:Energy_png}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Amino.pdf}
\caption{Amino acids in the presence of allosteric control. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 100 sets.}
\label{fig:Amino}
\end{figure}

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.94\textwidth]{./Figures/ClusterCAT161.pdf}
\caption{Normalized first-order and pairwise sensitivities of CAT production to rate constants. Dark pixels indicate the most sensitive indices. Dendrogram indicates high responders (dark gray background), medium responders (medium gray background), and low responders (light gray background).}
\label{fig:ClusterCAT}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=0.94\textwidth]{./Figures/ClusterCAT161.png}
\caption{Normalized first-order and pairwise sensitivities of CAT production to rate constants. Dark pixels indicate the most sensitive indices. Dendrogram indicates high responders (dark gray background), medium responders (medium gray background), and low responders (light gray background).}
\label{fig:ClusterCAT_png}
\end{figure}

@@
